---
title: "NURSERY SURVEY DATA"
author: "_"
date: "2024-11-14"
output:
  pdf_document: default
  html_document: default
---

# NURSERY SURVEY DATA - DATA ANALYSIS
--- Tree nurseries are integral to afforestation and reforestation initiatives as they provide the seedlings needed for successful restoration. This analysis focuses on survey data collected from various tree nurseries to assess their capacity to produce sufficient seedlings for ecological restoration projects.

The primary objective of this analysis is to understand the production capacity of nurseries across different counties. By examining variables such as the types of seedlings produced, nursery capacity, and other key indicators, we aim to identify regional strengths and gaps in the nurseries’ ability to meet restoration demands. Through the use of pivot tables, charts, and summary statistics, this analysis provides insights into the resources and capabilities available at each nursery and highlights areas for potential support or development.
## Installing packages for Analysis

```{r setup, include=FALSE}
# Set CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))
```



```{r}
install.packages("tidyverse")
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
```

## Importing the Data-set
```{r}
nursery_survey_data <- read_csv("nursery_survey_data.csv")
View(nursery_survey_data)
```

## Exploring the Data-set
```{r}
# checking the outlook of the data, the first 10 rows
head(nursery_survey_data, 10)
```

```{r}
# checking the attributes of the data
colnames(nursery_survey_data)
```

```{r}
#checking for missing values and NA's
summary(nursery_survey_data)
```

# Pivot Table for Organization, Respondent Status, and County
We will create a pivot table that summarizes the counts of respondent status (owner/employee) categorized by county (ADM1) and organization.
```{r}
# Summarize the counts of respondents categorized by status and county
pivot_table_status <- nursery_survey_data %>%
  group_by(ADM1, organization, respondent_status) %>%
  summarize(count = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = respondent_status,
    values_from = count,
    values_fill = list(count = 0)  # Fill missing values with 0
  )

# View the pivot table
print(pivot_table_status)
```


### Creating a Unique Identifier for Nurseries

```{r}
nursery_survey_data <- nursery_survey_data %>%
  mutate(nursery_id = paste(ADM3, nursery_location, sep = "_"))
```


# Summarizing the Nursery Capacity by County named as ADM1
To understand the capacity by county we group and summarize the data by County. This will give us insights into total and average seedling capacity in each county.To analyze the total and average seedling production capacity by county, we use group_by() and summarize() functions. Assume number_of_seedlings represents the capacity of each nursery.

```{r}
# Convert `number_of_seedlings` to numeric, handling any non-numeric values as NA
nursery_survey_data$number_of_seedlings <- as.numeric(nursery_survey_data$number_of_seedlings)

county_summary <- nursery_survey_data %>%
  group_by(ADM1) %>%
  summarize(
    total_seedlings = sum(number_of_seedlings, na.rm = TRUE),
    average_seedlings = mean(number_of_seedlings, na.rm = TRUE),
    nursery_count = n_distinct(nursery_id)  # Count unique nurseries
  )

# View summary
print(county_summary)
```

# Analyzing the nursery types and nursery purpose in each county
# Nursery Type Per County

```{r}
pivot_table_nursery <- nursery_survey_data %>%
  group_by(ADM1, nursery_type) %>%
  summarize(count = n(), .groups = "drop")

custom_colors <- c(
  "#4E79A7", "#59A14F", "#F28E2B", "#E15759", "#76B7B2", 
  "#FF9DA7", "#9C755F", "#B07AA1", "#F1C1D1", "#92C5DA", 
  "#B7C8F5"
)

ggplot(pivot_table_nursery, aes(x = ADM1, y = count, fill = ADM1)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~nursery_type) +  # Create separate plots for each nursery type
  scale_fill_manual(values = custom_colors) +  # Manually specify colors
  labs(
    title = "Nursery Type by County",
    x = "County",
    y = "Count",
    fill = "County"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"  # Remove the legend
  )
```
For Nursery type in each county, majority of them are group-led, followed by institutional, then individual and the least type is project-based

# Nursery Purpose
## Pivot tables for grouping the Nurseries by Purpose for each counties
```{r}
pivot_table_purpose <- nursery_survey_data %>%
  group_by(ADM1, nursery_purpose) %>%
  summarize(count = n(), .groups = "drop")  # Count the occurrences of each nursery purpose

# Step 2: Pivot the data to make it wider (create a column for each nursery_purpose)
pivot_table_wide_purpose <- pivot_table_purpose %>%
  pivot_wider(
    names_from = nursery_purpose,  # Creates columns for each nursery_purpose
    values_from = count,           # The count of each purpose in each county
    values_fill = list(count = 0)  # Fill missing values with 0
  )

# Step 3: View the pivot table
print(pivot_table_wide_purpose)
```


```{r}
purpose_count <- nursery_survey_data %>%
  group_by(nursery_purpose) %>%
  summarize(count = n(), .groups = "drop")  # Count the occurrences

# Step 2: Reorder nursery_purpose based on the count, from most to least frequent
purpose_count <- purpose_count %>%
  mutate(nursery_purpose = fct_reorder(nursery_purpose, count))

# Step 3: Create a bar graph with ggplot2
ggplot(purpose_count, aes(x = nursery_purpose, y = count, fill = nursery_purpose)) +
  geom_bar(stat = "identity") +  # Create a bar graph
  scale_fill_brewer(palette = "Set3") +  # Use a nice color palette
  labs(
    title = "Nursery Purpose Count",
    x = "Nursery Purpose",
    y = "Count",
    fill = "Nursery Purpose"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
From the analysis we can see that majority of the nurseries produce seeds for income forest management in-house needs while the least produce for income forest management in general not just their own in-house needs



# Finding the types of seedlings per county
```{r}
# Summarize the total counts of Indigenous and Exotic seedlings by county
pivot_table <- nursery_survey_data %>%
  group_by(ADM1) %>%
  summarize(
    count_indigenous = sum(Indigenous, na.rm = TRUE),
    count_exotic = sum(Exotic, na.rm = TRUE),
    total_count = count_indigenous + count_exotic,  # Calculate total count
    .groups = "drop"  # Avoid nesting the results
  )

# View the pivot table
print(pivot_table)
```


```{r}
pivot_table <- pivot_table %>%
  mutate(total_count = count_indigenous + count_exotic) %>%  # Calculate total seedlings per county
  arrange(desc(total_count)) %>%  # Arrange in descending order
  mutate(ADM1 = factor(ADM1, levels = unique(ADM1)))  # Reorder ADM1 based on total_count

# Step 2: Plot with reordered ADM1
ggplot(pivot_table, aes(x = ADM1)) +
  geom_bar(aes(y = count_indigenous, fill = "Indigenous"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = count_exotic, fill = "Exotic"), stat = "identity", position = "dodge") +
  labs(
    title = "Seedling Counts Across Counties",
    x = "County (ADM1)",
    y = "Count of Seedlings",
    fill = "Seedling Type"
  ) +
  scale_fill_manual(values = c("Indigenous" = "lightblue", "Exotic" = "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
We can conclude that Kiambu and Nakuru respectively produce most number of seeds and the least being Kajiado County

# Finding the registration status and registering Organizations for each County
## Rgistration Status
```{r}
# Step 1: Summarize and calculate proportions for registration status counts per county
pivot_table_reg_status <- nursery_survey_data %>%
  group_by(ADM1, nursery_registration_status) %>%
  summarise(count = n(), .groups = "drop") %>%
  complete(ADM1, nursery_registration_status, fill = list(count = 0)) %>%  # Fill missing with 0
  group_by(ADM1) %>%
  mutate(
    proportion = count / sum(count),  # Calculate proportions
    percentage = proportion * 100     # Calculate percentage for labeling
  )

# Step 2: Create the 100% stacked bar chart with percentage labels, filtering out near-0% values
ggplot(pivot_table_reg_status, aes(x = ADM1, y = proportion, fill = nursery_registration_status)) +
  geom_bar(stat = "identity", position = "fill") +  # 100% stacked bar chart
  geom_text(
    aes(label = ifelse(percentage >= 1, paste0(round(percentage, 1), "%"), "")),  # Show labels for values >= 1%
    position = position_fill(vjust = 0.5), size = 3
  ) +
  labs(
    title = "Registration Status Distribution per County (100% Stacked with Percentages)",
    x = "County",
    y = "Proportion",
    fill = "Registration Status"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("skyblue", "lightgreen", "coral", "purple", "orange")) +  # Define colors
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate county labels for readability
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = scales::percent)  # Show y-axis in percentage

```

## Registering Organization

```{r}
# Replace NA values with the string "NA" and summarize the data
pivot_table_registering_org <- nursery_survey_data %>%
  mutate(registering_org = ifelse(is.na(registering_org), "NA", registering_org)) %>%  # Replace NA with "NA"
  group_by(ADM1, registering_org) %>%
  summarise(count = n(), .groups = "drop")

# Ensure no empty slices (filter any counties with zero counts)
pivot_table_registering_org <- pivot_table_registering_org %>%
  filter(count > 0)

# Create a pie chart
ggplot(pivot_table_registering_org, aes(x = "", y = count, fill = registering_org)) +
  geom_bar(stat = "identity", width = 1) +  # Create the bar for the pie
  coord_polar(theta = "y") +  # Convert the bar chart to a pie chart
  facet_wrap(~ ADM1, scales = "free") +  # Facet by county (ADM1) with free scales
  labs(
    title = "Distribution of Registering Organizations by County",
    fill = "Registering Organization"
  ) +
  theme_void() +  # Clean up the chart for a more aesthetic look
  theme(axis.text.x = element_blank())  # Remove x-axis text
```

Among the counties, [County Name] stands out with a significantly higher number of registered organizations compared to the others, indicating a stronger presence of registered nurseries or a more active registration process in this region

# Analysing the Nursery Production Category per county

```{r}
# Create the pivot table: Count per nursery production category
pivot_table_nursery_production <- nursery_survey_data %>%
  group_by(ADM1, nursery_production_category) %>% 
  count() %>%  # Count the occurrences of each category per county
  spread(key = nursery_production_category, value = n, fill = 0)  # Reshape to wide format

# View the pivot table
print(pivot_table_nursery_production)
```

# Analysis of the Distance to Paved Road

```{r}
ggplot(nursery_survey_data, aes(x = reorder(ADM1, distance_to_paved_road, FUN = max), 
                                y = distance_to_paved_road, fill = ADM1)) +
  geom_bar(stat = "identity", position = "dodge") +  # Bar chart with the numeric distance
  labs(
    title = "Distance to Paved Road by County Type",
    x = "County Type",
    y = "Distance to Paved Road (in m)",
    fill = "County Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
  scale_fill_brewer(palette = "Set3")  # Use a readable color palette for fill
```


```{r}
# Create the bar chart with descending order for the bars based on distance_to_water
ggplot(nursery_survey_data, aes(x = reorder(ADM1, distance_to_water, FUN = max), y = distance_to_water, fill = ADM1)) +
  geom_bar(stat = "identity", position = "dodge") +  # Bar chart with the numeric distance
  labs(
    title = "Distance to Water by County Type",
    x = "County Type",
    y = "Distance to Water (in m)",
    fill = "County Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
  scale_fill_brewer(palette = "Set3")  # Use a readable color palette for fill
```

# Identifying the total number of employeers Per County
```{r}
# Summarize the data and calculate the total employees per county
employee_summary <- nursery_survey_data %>%
  group_by(ADM1) %>%
  summarize(
    Males_Total = sum(males_total, na.rm = TRUE),  # Total males
    Female_Total = sum(females_total, na.rm = TRUE),  # Total females
    Total_Employees = sum(males_total, na.rm = TRUE) + sum(females_total, na.rm = TRUE),  # Sum of males and females
    .groups = "drop"  # To avoid nesting
  )

# View the summary table with each county in its own row
print(employee_summary)
```

# Checking for which Counties take part in Seed Collection

```{r}
# Create a pivot table for Seed_Collection status per county (ADM1)
pivot_table_seed_collection <- nursery_survey_data %>%
  filter(!is.na(Seed_Collection)) %>%  # Remove NAs from Seed_Collection column
  group_by(ADM1, Seed_Collection) %>%  # Group by County and Seed Collection Status
  summarize(Count = n(), .groups = "drop") %>%  # Count occurrences
  pivot_wider(names_from = Seed_Collection, values_from = Count, values_fill = list(Count = 0))  # Pivot the table

# View the pivot table
print(pivot_table_seed_collection)
```

# Checking for Counies that Practise NUrsery Technques

```{r}
# Create a pivot table for Nursery_Techniques per county (ADM1)
pivot_table_nursery_techniques <- nursery_survey_data %>%
  filter(!is.na(Nursery_Techniques)) %>%  # Remove NAs from Nursery_Techniques column
  group_by(ADM1, Nursery_Techniques) %>%  # Group by County and Nursery Techniques (YES/NO)
  summarize(Count = n(), .groups = "drop") %>%  # Count occurrences
  pivot_wider(names_from = Nursery_Techniques, values_from = Count, values_fill = list(Count = 0))  # Pivot the table

# View the pivot table
print(pivot_table_nursery_techniques)
```
It is evident that Kjiado County do not practice nursery techniques and also do not take part in the seed collection. All the rest take part in Nursery techniques apart from some but for some from 16 from Nakuru and 5 from Nyandarua.

# Seed sales per County
```{r}
# Create a pivot table to count NO for Seed_Sales per nursery_id (case-insensitive)
pivot_table_no_seed_sales_nursery <- nursery_survey_data %>%
  filter(tolower(Seed_Sales) == "no") %>%  # Convert Seed_Sales to lowercase and filter for "no"
  group_by(nursery_id) %>%  # Group by nursery_id
  summarize(Count = n(), .groups = "drop")  # Count the occurrences of "no" for each nursery_id

# View the pivot table
print(pivot_table_no_seed_sales_nursery)
```

The above is a list of all nurseries that do not take part in Seed Sales because they cannot meet demands or are yet to make sales, they don't sale since their goal is restoration and conservation, They have partnered with KFS to plant in their forests and they support them in marketing the seedlings,they have the reliable project market from the ITF or their in-house market demand is very promising 

# Nursery Infrasture Appropriateness Per County

```{r}
# Data preparation: Count nursery_infrastructure_appropriateness responses per county
infrastructure_data <- nursery_survey_data %>%
  filter(!is.na(Nursery_Infrastructure_Appropriateness)) %>%  # Remove NA values
  group_by(ADM1, Nursery_Infrastructure_Appropriateness) %>%
  summarize(count = n(), .groups = "drop")  # Count responses for each county and appropriateness status

# Define custom shades of blue for each infrastructure appropriateness level
blue_shades <- c("poor" = "#bdd7e7", "fair" = "#4169e1", "good" = "blue", "very_good" = "#00008B")

# Create the stacked bar chart
ggplot(infrastructure_data, aes(x = ADM1, y = count, fill = Nursery_Infrastructure_Appropriateness)) +
  geom_bar(stat = "identity") +  # Default stacking for bars
  labs(
    title = "Nursery Infrastructure Appropriateness per County",
    x = "County",
    y = "Count",
    fill = "Infrastructure Appropriateness"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
  scale_fill_manual(values = blue_shades)  # Apply custom shades of blue
```
From the graph we discover that Kiambu County has the most very_good infrastructure appropriateness followed by Nakuru. Nakuru has a ratio of poor infrastructure whereas Kajiado and West-Pokot have complete fair and poor infrastructure. Laikipia, Meru and Narok have good infrastructures. The rest that is Baringo, Elgeyo Marakwet, Makueni and Nyandarua have a share of both good and fair infrastructures. Kiambu has a mix of both very good and fair while Nakuru seems to have signifacant ratios of all 4 categories


# Materials Satisfaction Per
```{r}
# Data preparation: Count nursery materials satisfaction responses per county
satisfaction_data <- nursery_survey_data %>%
  filter(!is.na(Nursery_Materials_Satisfaction)) %>%  # Remove NA values
  group_by(ADM1, Nursery_Materials_Satisfaction) %>%
  summarize(count = n(), .groups = "drop")  # Count responses for each county and satisfaction level

# Define custom shades of green for each satisfaction level
green_shades <- c("not_satisfied" = "#a1d99b", "somehow_satisfied" = "#006d2c", 
                  "satisfied" = "#31a354", "very_satisfied" = "black")

# Create the stacked bar chart
ggplot(satisfaction_data, aes(x = ADM1, y = count, fill = Nursery_Materials_Satisfaction)) +
  geom_bar(stat = "identity") +  # Default stacking for bars
  labs(
    title = "Nursery Materials Satisfaction per County",
    x = "County",
    y = "Count",
    fill = "Satisfaction Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
  scale_fill_manual(values = green_shades)  # Apply custom shades of green
```

From the bars, we can identify that Kiambu has a ratio of all satisfactions but the majority are unsatisfied with the nursery materials present. For the rest of the counties there is some level of satisfied apart from Kajiado, Narok and Nyandarua that incline more to somehow satisfied and not satisfied

# SPECIES ANALYSIS
## Analysing the number of distinct Species present in each County

```{r}
# Create a pivot table for distinct species count per county
species_count_per_county <- nursery_survey_data %>%
  group_by(ADM1) %>%  # Group by county
  summarize(distinct_species_count = n_distinct(species_name, na.rm = TRUE), .groups = "drop")  # Count distinct species

# View the pivot table
print(species_count_per_county)
```

## Number of Seedlings available Per County
```{r}
# Create a pivot table for the total number of seedlings per county and arrange in descending order
seedlings_count_per_county <- nursery_survey_data %>%
  group_by(ADM1) %>%  # Group by county
  summarize(total_seedlings = sum(number_of_seedlings, na.rm = TRUE), .groups = "drop") %>%  # Sum of seedlings per county
  arrange(desc(total_seedlings))  # Arrange in descending order

# View the pivot table
print(seedlings_count_per_county)
```

## Method of Propagation used in each County
```{r}
# Create a pivot table counting different methods of propagation per county
method_of_propagation_per_county <- nursery_survey_data %>%
  group_by(ADM1) %>%
  summarize(
    seeds = sum(method_of_propagation == "seeds", na.rm = TRUE),
    vegetative = sum(method_of_propagation == "vegetative", na.rm = TRUE),
    wildings = sum(method_of_propagation == "wildings", na.rm = TRUE),
    wildings_and_seeds = sum(method_of_propagation == "wildings and seeds", na.rm = TRUE),
    NA_count = sum(is.na(method_of_propagation))
  ) %>%
  ungroup()  # To ensure that grouping is dropped after summarizing

# View the pivot table
print(method_of_propagation_per_county)
```

## checking for Collection Sourecs for each County
```{r}
# Create the pivot table for own_collection_source per county
pivot_table_own_collection <- nursery_survey_data %>%
  group_by(ADM1, own_collection_source) %>%
  summarize(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = own_collection_source, values_from = count, values_fill = list(count = 0))

# View the pivot table
print(pivot_table_own_collection)
```

# ANALYSING SELLERS
## Checking whether the sellers attended training offered
```{r}
# Create the pivot table for training_attended per county with "No" changed to "no"
pivot_table_training_attended <- nursery_survey_data %>%
  mutate(training_attended = tolower(training_attended)) %>%  # Convert "No" to "no"
  group_by(ADM1, training_attended) %>%
  summarize(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = training_attended, values_from = count, values_fill = list(count = 0))

# View the pivot table
print(pivot_table_training_attended)
```

## Analysis from the sellers whether the training was helpful for them in relation to the theme
```{r}
# Create a pivot table that categorizes `training_theme` using `training_helpful`
pivot_table_training_theme <- nursery_survey_data %>%
  group_by(training_theme, ADM1, training_helpful) %>%  # Group by both training theme and helpfulness
  summarize(count = n(), .groups = "drop") %>%   # Count the occurrences
  pivot_wider(names_from = training_helpful, values_from = count, values_fill = list(count = 0))

# View the pivot table
print(pivot_table_training_theme)
```

From the table we can see that 26 themes were involved the training sessions, 21 received no training there were NA's from the data showing the training sessions were not availed as expected in all counties. Only 2 organizations provided the Training, that is, ICRAF and Green Belt Movement


# MARKET ANALYSIS
## Established Markets Per County
```{r}
# Change "No" to lowercase "no" in the Established_Market column
nursery_survey_data <- nursery_survey_data %>%
  mutate(Established_Market = str_to_lower(Established_Market))

# Summarize data for Established markets per county
established_markets_summary <- nursery_survey_data %>%
  group_by(ADM1, Established_Market) %>%
  summarize(count = n(), .groups = "drop")

# Create a stacked bar chart and add counts on top of the bars
ggplot(established_markets_summary, aes(x = ADM1, y = count, fill = Established_Market)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bar chart
  geom_text(aes(label = count), position = position_stack(vjust = 0.5), size = 3.5) +  # Add count labels inside the bars
  labs(
    title = "Established Markets Per County",
    x = "County",
    y = "Count of Established Markets",
    fill = "Established Markets"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Identifying the main customers from Each of the Counties
```{r}
# Summarize the count of Main_Customers per county
main_customers_summary <- nursery_survey_data %>%
  group_by(ADM1, Main_Customers) %>%
  summarize(count = n(), .groups = "drop")

# Create a 100% stacked bar chart for Main_Customers per county
ggplot(main_customers_summary, aes(x = ADM1, y = count, fill = Main_Customers)) +
  geom_bar(stat = "identity", position = "fill") +  # Use "fill" for 100% stacked bar chart
  geom_text(
    aes(label = scales::percent(count / sum(count), accuracy = 1)),
    position = position_fill(vjust = 0.5),  # Center text within each bar
    size = 3
  ) +
  labs(
    title = "Main Customers per County",
    x = "County",
    y = "Proportion of Main Customers",
    fill = "Main Customers"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Species Preferred Per County
```{r}
# Summarize the count of Main_Customers per county
species_preferred_summary <- nursery_survey_data %>%
  group_by(ADM1, Species_Preferred) %>%
  summarize(count = n(), .groups = "drop")

# Create a 100% stacked bar chart for Main_Customers per county
ggplot(species_preferred_summary, aes(x = ADM1, y = count, fill = Species_Preferred)) +
  geom_bar(stat = "identity", position = "fill") +  # Use "fill" for 100% stacked bar chart
  geom_text(
    aes(label = scales::percent(count / sum(count), accuracy = 1)),
    position = position_fill(vjust = 0.5),  # Position text inside bars
    size = 3
  ) +
  labs(
    title = "Species Preferred per County",
    x = "County",
    y = "Proportion of Species Preferred",
    fill = "Species Preferred"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Marketing Medium
```{r}
# Summarize the count of Marketing_Medium per County
marketing_medium_summary <- nursery_survey_data %>%
  group_by(ADM1, Marketing_Medium) %>%
  summarize(count = n(), .groups = "drop")

# Create a 100% stacked bar chart for Marketing_Medium per County
ggplot(marketing_medium_summary, aes(x = ADM1, y = count, fill = Marketing_Medium)) +
  geom_bar(stat = "identity", position = "fill") +  # Use "fill" for 100% stacked bar chart
  geom_text(
    aes(label = scales::percent(count / sum(count), accuracy = 1)),
    position = position_fill(vjust = 0.5),  # Center the text within each bar
    size = 3
  ) +
  labs(
    title = "Marketing Medium per County",
    x = "County",
    y = "Proportion of Marketing Mediums",
    fill = "Marketing Medium"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Aalysis of The Type of Constraints in the Market Place
```{r}
# Summarize the count of Constraints_To_Reaching_Customers per county
constraints_summary <- nursery_survey_data %>%
  group_by(ADM1, Constraints_To_Reaching_Customers) %>%
  summarize(count = n(), .groups = "drop")

# Calculate the percentage for each constraint per county
constraints_summary <- constraints_summary %>%
  group_by(ADM1) %>%
  mutate(percentage = count / sum(count) * 100)

# Plot pie charts for each county
ggplot(constraints_summary, aes(x = "", y = percentage, fill = Constraints_To_Reaching_Customers)) +
  geom_bar(stat = "identity", width = 1) +  # Create the bar chart
  coord_polar(theta = "y") +  # Convert to pie chart
  facet_wrap(~ ADM1) +  # Create a separate pie chart for each county
  labs(
    title = "Distribution of Constraints to Reaching Customers per County",
    fill = "Constraints"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),  # Hide x-axis text
        axis.ticks = element_blank(),  # Remove axis ticks
        panel.grid = element_blank())  # Remove grid lines
```

## Market Information Source
```{r}

# Step 1: Summarize the count of nursery_mgt_information_source per county
mgt_info_summary <- nursery_survey_data %>%
  group_by(ADM1, nursery_mgt_information_source) %>%
  summarize(count = n(), .groups = "drop")

# Step 2: Pivot the data to a wide format to display actual counts per county
pivot_table <- mgt_info_summary %>%
  spread(key = nursery_mgt_information_source, value = count, fill = 0)  # Pivoting the data

# Step 3: Display the pivot table using kable
kable(pivot_table, caption = "Distribution of Nursery Management Information Sources per County") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), position = "center") %>%
  add_header_above(c(" " = 1, "Information Sources" = ncol(pivot_table) - 1)) %>%
  row_spec(0, bold = TRUE)  # Bold the header row
```


# NURSERY MANAGEMENT ANALYSIS
## Checking if the Nurseries in the particular counties plan on scaling up
```{r}
# Step 1: Summarize the count of scale_up_constraints across all counties
constraints_summary_global <- nursery_survey_data %>%
  group_by(scale_up_constraints) %>%
  summarize(count = n(), .groups = "drop")

# Step 2: Calculate the overall percentage for each scale_up_constraints
constraints_summary_global <- constraints_summary_global %>%
  mutate(percentage = count / sum(count) * 100)

# Step 3: Create a pivot table to show scale_up_constraints and their overall percentage
pivot_table_constraints <- constraints_summary_global %>%
  select(scale_up_constraints, percentage) %>%
  pivot_wider(names_from = scale_up_constraints, values_from = percentage)

# View the pivot table
print(pivot_table_constraints)
```

## Competitors per County
```{r}
# Summarize competitors_list by county
competitors_summary <- nursery_survey_data %>%
  group_by(ADM1, competitors_list) %>%
  summarize(count = n(), .groups = "drop") %>%
  spread(key = competitors_list, value = count, fill = 0)  # Pivot table format

# View the pivot table
print(competitors_summary)
```

# Training Areas Required

```{r}
# Step 1: Summarize the data by county and training areas required
training_summary <- nursery_survey_data %>%
  group_by(ADM1, training_areas_required) %>%
  summarize(count = n(), .groups = "drop")

# Step 2: Create a pivot table to show counts of training areas required by county
pivot_table <- training_summary %>%
  spread(key = training_areas_required, value = count, fill = 0)  # Pivoting the table

# Step 3: Display the pivot table using kable and kableExtra for formatting
kable(pivot_table, caption = "Training Areas Required per County") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), position = "center") %>%
  add_header_above(c(" " = 1, "Training Areas" = ncol(pivot_table)-1)) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE)  # Make header row bold
```





